Physcraper: A Python package for continually updated gene trees

### 2020-05-19 phylota-overview.Rmd
starting to outline discussion

@@

@@

1. Other miscellaneous uses of phylota:

@@

>>>>

    - A review on online plant databases aiming to "provide recommendations for current
    information managers and developers concerning the user interface and experience;
    and to provide a picture about the possible directions to take for those in charge
    of the creation of information at all levels". They cite phylota as a tool allowing
    researchers "to acces equally and globally, without travel, a [phylogenetic?] model
    of plants at the kingdom level".
    - A study presenting a tool to asses gene sequence quality for automatic
    construction of databases [@meng2012gsqct], as well as their parallelized version using MapReduce
    [@meng2012cloud], cite phylota (along with @yong2010screening) as a tool that relies on sequence
    similarity (BLAST) and not taxon name annotations in the database, for mining
    large numbers of taxa or loci, without making any control on the quality of the
    sequencing. On the same line
    - a paper aiming to establish an online information system for the legumes and
    to outline "best practices for development of a legume portal to enable data
    sharing and a better understanding of what data are available, missing, or erroneous,
    and ultimately facilitate cross-analyses and collaboration within the legume-systematics
    community and with other stakeholders" [@bruneau2019towards], cites phylota (along with supersmart and pyphlawd) as a
    "pipeline for large-scale retrieval of GenBank data of particular taxa or clades".
    In their Table 1, they also list phylota as a potential data source for developing a legume portal.

@@

    - @kolmann2017dna uses phylota to download all 1691 co1 sequences belonging to
    the order Carchariniformes, to place phylogenetically DNA samples obtained from
    fish markets.
    - A phylogeny of 329 bird species in the Tyrannidae (77% of the species in the
      family) [@gomez2020speciation]

@@

1. When phylota was used to extract full trees (not only DNA data sets or markers):

@@

    - @deepak2013extracting uses a sample of phylota trees to test their method to
    remove conflict from MUL-trees (short for multi-labeled trees), that is, phylogenetic
    trees with two or more leaves sharing a label, e.g., a species name, which can
    imply multiple conflicting phylogenetic relationships for the same set of taxa.

### 2020-05-19 physcraper_ms.Rmd
starting to outline discussion

Phylogenies are important.
Generating phylogenies is not easy.
The process of phylogenetic reconstruction implies many steps that can be generalized to the following:
1. Obtention of molecular or morphological character data -- get DNA from some organisms and sequence it, or get it from an online repository, such as GenBank.
1. Assemble a hypothesis of homology -- Create a matrix of your character data, by aligning the sequences, in the case of molecular data.
1. Analyse this hypothesis of homology to infer phylogenetic relationships among the organisms you are studying -- Use different available programs to infer molecular evolution, trees and times of divergence.
1. Discuss the inferred relationships in the context of previous hypothesis, the biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*

@@

```
physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
```

# Tools on a similar track

Tools that do similar things at different levels:

@@

>>>>

Phylogenies are important.
Generating phylogenies is not easy.
The process of phylogenetic reconstruction implies many steps that can be generalized to the following:

1. Obtention of molecular or morphological character data -- get DNA from some organisms
and sequence it, or get it from an online repository, such as GenBank.   
1. Assemble a hypothesis of homology -- Create a matrix of your character data, by
aligning the sequences, in the case of molecular data.
1. Analyse this hypothesis of homology to infer phylogenetic relationships among
the organisms you are studying -- Use different available programs to infer molecular
evolution, trees and times of divergence.
1. Discuss the inferred relationships in the context of previous hypothesis, the
biology and biogeography of the organisms, etc. -- Answer the question, *is this phylogenetic solution fair/reasonable?*

@@

```
physcraper_run.py -s pg_2573 -t tree5959 -tb -db ~/branchinecta/local_blast_db/ -o pg_2573
```
# Discussion

Mention statistics provided by PhyloExplorer [@ranwez2009phyloexplorer]

## Tools that do similar things at different levels

PhyloFinder [@chen2008phylofinder] - a search engine for phylogenetic databases using
trees from TreeBASE - MEH

@@

PhyloExplorer [@ranwez2009phyloexplorer] - a python and MySQL based website to facilitate
assessment and management of phylogenetic tree collections.
It provides "statistics describing the collection,
correcting invalid taxon names, extracting taxonomically relevant parts of the collection
using a dedicated query language, and identifying related trees in the TreeBASE database".
